---
title: "Company reports over time"
format: html
---

## Purpose

I want to create a graph of the company reports over time. I want to have year on the x-axis and the name of the company on the y-axis. I want to have a dot for each year-company that shows the number of reports that company has in that year.

Some firms have joined over time, so there may be instances of multiple reports per year, or special reports that are released with a specific purpose.

```{r}
library(tidyverse)

df <- readxl::read_excel(here::here("data", "temp", "raw_pdf_filenames_edit.xlsx"))

df <- df %>%
    filter(!is.na(company_name)) %>%
    mutate(
        # year is a column where we extract the first instance of four digits in the filename
        year = str_extract(filename, "\\d{4}"),
        # we convert the year to a number
        year = as.numeric(year)
    )

firms_to_plot <- df %>%
    count(company_name) %>%
    filter(n > 30) %>%
    # slice_sample(n = 20)  %>%
    select(company_name) %>%
    pull()

firms_to_plot %>%
    as_tibble() %>%
    rename(company_name = value) %>%
    write_csv(here::here("data", "temp", "firms_to_plot.csv"))

df_order <- df %>%
    filter(company_name %in% firms_to_plot) %>%
    group_by(company_name) %>%
    mutate(mean_year = mean(year, na.rm = T)) %>%
    select(company_name, mean_year) %>%
    ungroup() %>%
    distinct()

df %>%
    count(year, company_name) %>%
    filter(company_name %in% firms_to_plot) %>%
    inner_join(df_order, by = "company_name") %>%
    mutate(company_name = fct_reorder(company_name, mean_year)) %>%
    ggplot(aes(x = year, y = company_name, fill = n)) +
    geom_tile() +
    theme_minimal() +
    theme(legend.position = "bottom") +
    scale_fill_distiller(palette = "Spectral", breaks = c(1, 2, 3, 4)) +
    guides(fill = guide_legend(title.position = "top")) +
    labs(
        title = "Company reports over time",
        x = "Year",
        y = "Company",
        fill = "Number of reports in one year"
    )

ggsave(here::here("images", "company_reports_over_time.png"), width = 10, height = 10, dpi = 300)
```



```{r}
library(tibble)

board_members <- tibble::tibble(
    year = c(
        1901, 1901, 1901, 1901, 1901,
        1911, 1911, 1911, 1911, 1911, 1911, 1911,
        1921, 1921, 1921, 1921, 1921, 1921, 1921,
        1931, 1931, 1931, 1931, 1931, 1931,
        1941, 1941, 1941, 1941, 1941, 1941,
        1951, 1951, 1951, 1951, 1951, 1951,
        1961, 1961, 1961, 1961, 1961, 1961, 1961,
        1971, 1971, 1971, 1971, 1971, 1971, 1971,
        1981, 1981, 1981, 1981, 1981, 1981, 1981, 1981, 1981, 1981, 1981, 1981, 1981, 1981, 1981
    ),
    name = c(
        "O. Fr. Wijkman", "V. Åkerman", "G. A. Granström", "Sven Gyllensvärd", "G. Wenström",
        "P. T. Berg", "John Fredholm", "O. Wallenberg", "John Karlsson", "J. L. la Cour", "H. E. Henke", "J. S. Edström",
        "O. Wallenberg", "W:m Henriques", "Sten Ankarcrona", "John Karlsson", "J. S. Edström", "P. T. Berg", "Arthur Lindén",
        "Sten Ankarcrona", "Axel F. Enström", "M. Wallenberg jr.", "Axel Bergengren", "J. S. Edström", "Arthur Lindén",
        "J. Sigfrid Edström", "Axel F. Enström", "M. Wallenberg jr.", "Axel Bergengren", "Walter Murray", "Arthur Lindén",
        "M. Wallenberg", "Thorsten Ericson", "Ruben Rausing", "Ragnar Söderberg", "Dan-Axel Broström", "Åke T. Vrethem",
        "M. Wallenberg", "Ragnar Söderberg", "Dan-Axel Broström", "Curt Nicolin", "Ruben Rausing", "Åke T. Vrethem", "Karl Nilsson",
        "Marcus Wallenberg", "Ragnar Söderberg", "Åke T. Vrethem", "Dan-Axel Broström", "Curt Nicolin", "Alde Nilsson", "Halvard Liander",
        "Marcus Wallenberg", "Curt Nicolin", "Åke T. Vrethem", "Peter Wallenberg", "Erich Aas", "Gunnar Andersson", "Percy Barnevik", "Torsten Lindström", "Sven Söderberg", "Folke Westerberg", "Carl Andersson", "Björn Berggren", "Alde Nilsson", "Christer Salén", "Carl-Bertil Winqvist"
    ),
    index = c(
        1, 2, 3, 4, 5,
        6, 7, 8, 9, 10, 11, 12,
        8, 13, 14, 9, 12, 6, 15,
        14, 16, 17, 18, 12, 15,
        12, 16, 17, 18, 19, 15,
        17, 20, 21, 22, 23, 24,
        17, 22, 23, 25, 21, 24, 26,
        17, 22, 24, 23, 25, 27, 28,
        17, 25, 24, 29, 30, 31, 32, 33, 34, 35, 36, 37, 27, 38, 39
    )
)

board_members

display_names <- board_members %>%
    select(index, name) %>%
    group_by(index) %>%
    summarise(name = first(name)) %>%
    ungroup() %>%
    rename(name_display = name)

board_members %>%
    inner_join(display_names, by = "index") %>%
    select(year, name_display) %>%
    group_by(name_display) %>%
    mutate(
        mid_year = mean(year, na.rm = T),
        board_count = n()
    ) %>%
    ungroup() %>%
    mutate(name_display = fct_reorder(name_display, mid_year)) %>%
    filter(board_count > 1) %>%
    ggplot(aes(x = year, y = name_display, colour = factor(board_count))) +
    geom_point() +
    geom_line() +
    scale_colour_brewer(palette = "Dark2") +
    labs(colour = "Number of decades on ASEA board") +
    theme(legend.position = "bottom") 

ggsave(here::here("images", "board_members_over_time.png"), width = 10, height = 10, dpi = 300)

```