---
title: "Company reports over time"
format: html
---

## Purpose

I want to create a graph of the company reports over time. I want to have year on the x-axis and the name of the company on the y-axis. I want to have a dot for each year-company that shows the number of reports that company has in that year.

Some firms have joined over time, so there may be instances of multiple reports per year, or special reports that are released with a specific purpose.

```{r}
library(tidyverse)

df <- readxl::read_excel(here::here("data", "temp", "raw_pdf_filenames_edit.xlsx"))

df <- df %>%
    filter(!is.na(company_name)) %>%
    mutate(
        # year is a column where we extract the first instance of four digits in the filename
        year = str_extract(filename, "\\d{4}"),
        # we convert the year to a number
        year = as.numeric(year)
    )

firms_to_plot <- df %>%
    count(company_name) %>%
    filter(n > 30) %>%
    # slice_sample(n = 20)  %>%
    select(company_name) %>%
    pull()

firms_to_plot %>%
    as_tibble() %>%
    rename(company_name = value) %>%
    write_csv(here::here("data", "temp", "firms_to_plot.csv"))

df_order <- df %>%
    filter(company_name %in% firms_to_plot) %>%
    group_by(company_name) %>%
    mutate(mean_year = mean(year, na.rm = T)) %>%
    select(company_name, mean_year) %>%
    ungroup() %>%
    distinct()

df %>%
    count(year, company_name) %>%
    filter(company_name %in% firms_to_plot) %>%
    inner_join(df_order, by = "company_name") %>%
    mutate(company_name = fct_reorder(company_name, mean_year)) %>%
    ggplot(aes(x = year, y = company_name, fill = n)) +
    geom_tile() +
    theme_minimal() +
    theme(legend.position = "bottom") +
    scale_fill_distiller(palette = "Spectral", breaks = c(1, 2, 3, 4)) +
    guides(fill = guide_legend(title.position = "top")) +
    labs(
        title = "Company reports over time",
        x = "Year",
        y = "Company",
        fill = "Number of reports in one year"
    )

ggsave(here::here("images", "company_reports_over_time.png"), width = 10, height = 10, dpi = 300)
```



```{r}
library(tibble)

board_members <- tibble::tibble(
    year = c(
        1901, 1901, 1901, 1901, 1901,
        1911, 1911, 1911, 1911, 1911, 1911, 1911,
        1921, 1921, 1921, 1921, 1921, 1921, 1921,
        1931, 1931, 1931, 1931, 1931, 1931,
        1941, 1941, 1941, 1941, 1941, 1941,
        1951, 1951, 1951, 1951, 1951, 1951,
        1961, 1961, 1961, 1961, 1961, 1961, 1961,
        1971, 1971, 1971, 1971, 1971, 1971, 1971,
        1981, 1981, 1981, 1981, 1981, 1981, 1981, 1981, 1981, 1981, 1981, 1981, 1981, 1981, 1981
    ),
    name = c(
        "O. Fr. Wijkman", "V. Åkerman", "G. A. Granström", "Sven Gyllensvärd", "G. Wenström",
        "P. T. Berg", "John Fredholm", "O. Wallenberg", "John Karlsson", "J. L. la Cour", "H. E. Henke", "J. S. Edström",
        "O. Wallenberg", "W:m Henriques", "Sten Ankarcrona", "John Karlsson", "J. S. Edström", "P. T. Berg", "Arthur Lindén",
        "Sten Ankarcrona", "Axel F. Enström", "M. Wallenberg jr.", "Axel Bergengren", "J. S. Edström", "Arthur Lindén",
        "J. Sigfrid Edström", "Axel F. Enström", "M. Wallenberg jr.", "Axel Bergengren", "Walter Murray", "Arthur Lindén",
        "M. Wallenberg", "Thorsten Ericson", "Ruben Rausing", "Ragnar Söderberg", "Dan-Axel Broström", "Åke T. Vrethem",
        "M. Wallenberg", "Ragnar Söderberg", "Dan-Axel Broström", "Curt Nicolin", "Ruben Rausing", "Åke T. Vrethem", "Karl Nilsson",
        "Marcus Wallenberg", "Ragnar Söderberg", "Åke T. Vrethem", "Dan-Axel Broström", "Curt Nicolin", "Alde Nilsson", "Halvard Liander",
        "Marcus Wallenberg", "Curt Nicolin", "Åke T. Vrethem", "Peter Wallenberg", "Erich Aas", "Gunnar Andersson", "Percy Barnevik", "Torsten Lindström", "Sven Söderberg", "Folke Westerberg", "Carl Andersson", "Björn Berggren", "Alde Nilsson", "Christer Salén", "Carl-Bertil Winqvist"
    ),
    index = c(
        1, 2, 3, 4, 5,
        6, 7, 8, 9, 10, 11, 12,
        8, 13, 14, 9, 12, 6, 15,
        14, 16, 17, 18, 12, 15,
        12, 16, 17, 18, 19, 15,
        17, 20, 21, 22, 23, 24,
        17, 22, 23, 25, 21, 24, 26,
        17, 22, 24, 23, 25, 27, 28,
        17, 25, 24, 29, 30, 31, 32, 33, 34, 35, 36, 37, 27, 38, 39
    )
)

board_members

display_names <- board_members %>%
    select(index, name) %>%
    group_by(index) %>%
    summarise(name = first(name)) %>%
    ungroup() %>%
    rename(name_display = name)

board_members %>%
    inner_join(display_names, by = "index") %>%
    select(year, name_display) %>%
    group_by(name_display) %>%
    mutate(
        mid_year = mean(year, na.rm = T),
        board_count = n()
    ) %>%
    ungroup() %>%
    mutate(name_display = fct_reorder(name_display, mid_year)) %>%
    filter(board_count > 1) %>%
    ggplot(aes(x = year, y = name_display, colour = factor(board_count))) +
    geom_point() +
    geom_line() +
    scale_colour_brewer(palette = "Dark2") +
    labs(colour = "Number of decades on ASEA board") +
    theme(legend.position = "bottom") 

ggsave(here::here("images", "board_members_over_time.png"), width = 10, height = 10, dpi = 300)

```




### Read in data from Excel about the share of technical and non-technical board members


```{r}
library(readxl)
library(tidyverse)

df <- readxl::read_excel(here::here("data/director_education_summaries/director_education_shares.xlsx"))

df %>% select(company, year, share_technical, share_business) %>%
    pivot_longer(cols = c(share_technical, share_business), names_to = "type", values_to = "value") %>%
    ggplot(aes(x = year, y = value, colour = type)) +
    geom_line() +
    facet_wrap(~ company) +
    theme(legend.position = "bottom") +
    labs(colour = "Type of education")


df %>% select(company, year, share_non_swedish, share_usa) %>%
    pivot_longer(cols = c(share_non_swedish, share_usa), names_to = "type", values_to = "value") %>%
    ggplot(aes(x = year, y = value, colour = type)) +
    geom_line() +
    facet_wrap(~ company) +
    theme(legend.position = "bottom") +
    labs(colour = "Type of education")
```

Contrast ASEA and Electrolux


```{r}
df %>%
    select(company, year, share_technical, share_business) %>%
    filter(company %in% c("ASEA", "Electrolux")) %>%
    pivot_longer(cols = c(share_technical, share_business), names_to = "type", values_to = "value") %>%
    ggplot(aes(x = year, y = value, colour = type, lty = type)) +
    geom_line() +
    facet_wrap(~company) +
    theme(legend.position = "bottom") +
    labs(
        colour = "Type of education of board members",
        lty = "Type of education of board members"
    ) +
    scale_y_continuous(labels = scales::percent_format(scale = 100)) +
    scale_colour_manual(values = c("blue", "red"), labels = c("Technical", "Business")) +
    labs(
        title = "Share of technical and business education on board over time",
        x = "Year",
        y = "Share of board members with specific education"
    )


df %>%
    select(company, year, share_non_swedish, share_usa) %>%
    filter(company %in% c("ASEA", "Electrolux")) %>%
    pivot_longer(cols = c(share_non_swedish, share_usa), names_to = "type", values_to = "value") %>%
    ggplot(aes(x = year, y = value, colour = type)) +
    geom_line() +
    facet_wrap(~company) +
    theme(legend.position = "bottom") +
    labs(colour = "Type of foreign experience of board members") +
    scale_y_continuous(labels = scales::percent_format(scale = 100)) +
    scale_colour_manual(values = c("blue", "red"), labels = c("Outside of Sweden", "USA")) 
```



Plot of share of firms with technical education on board over time, and share of firms with business education on board over time

```{r}
df %>%
    filter(between(year, 1900, 1990)) %>%
    group_by(year) %>%
    summarise(
        share_technical = mean(share_technical),
        share_business = mean(share_business)
    ) %>%
    ggplot(aes(x = year)) +
    geom_line(aes(y = share_technical, colour = "Technical")) +
    geom_line(aes(y = share_business, colour = "Business")) +
    theme(legend.position = "bottom") +
    labs(colour = "Type of education of board members") +
    scale_y_continuous(labels = scales::percent_format(scale = 100)) +
    scale_colour_manual(values = c("blue", "red"), labels = c("Technical", "Business"))

```


Do the shares of technical education increase over time?

```{r}
df_directors <- readxl::read_excel("data/directors_education_data_deduplicated.xlsx")

df_directors %>%
    # extract 4 digits from the birth_date column
    mutate(
        birth_year = str_extract(birth_date, "\\d{4}"),
        birth_year = parse_number(birth_year)
    ) %>%
    mutate(
        # create variable to smooth out the data, where birthyear is grouped by 5 years
        birth_year_5 = floor(birth_year / 3) * 3,
    ) %>%
    group_by(birth_year_5) %>%
    summarise(
        mean_technical = mean(has_technical_education, na.rm = T),
        mean_business = mean(has_business_education, na.rm = T),
    ) %>%
    filter(between(birth_year_5, 1860, 1950)) %>%
    ggplot(aes(x = birth_year_5)) +
    geom_line(aes(y = mean_technical, colour = "Technical"), lty = 1) +
    geom_line(aes(y = mean_business, colour = "Business"), lty = 2) +
    theme(legend.position = "bottom") +
    labs(
        colour = "Type of education of board members",
        x = "Year of director's birth",
        y = "Share of director's with specific education",
        caption = "Smoothing by 3 year rolling mean"
    ) +
    scale_y_continuous(labels = scales::percent_format(scale = 100)) +
    scale_colour_manual(values = c("blue", "red"), labels = c("Technical", "Business")) 

```




```{r}
df_directors %>%
    select(id, education) %>%
    mutate(institution = str_extract_all(education, "institution':\\s*'([^']+)")) %>%
    unnest(institution) %>%
    mutate(
        institution = str_remove_all(institution, "institution': '")
    ) %>%
    view()
```



## Now with the entire sample


```{r}
df_exper <- readxl::read_excel("data/all_education_data_deduplicated.xlsx")



df_exper_plot <- df_exper %>%
    # extract 4 digits from the birth_date column
    mutate(
        birth_year = str_extract(birth_date, "\\d{4}"),
        birth_year = parse_number(birth_year)
    ) %>%
    mutate(
        # create variable to smooth out the data, where birthyear is grouped by 5 years
        birth_year_5 = floor(birth_year / 3) * 3,
    ) %>%
    group_by(birth_year_5) %>%
    summarise(
        mean_technical = mean(has_technical_education, na.rm = T),
        mean_business = mean(has_business_education, na.rm = T),
    ) %>%
    filter(between(birth_year_5, 1860, 1950))



df_exper_plot %>%
    ggplot(aes(x = birth_year_5)) +
    geom_line(aes(y = mean_technical, colour = "Technical"), lty = 1) +
    geom_line(aes(y = mean_business, colour = "Business"), lty = 2) +
    theme(legend.position = "bottom") +
    labs(
        colour = "Type of education of board members",
        x = "Year of director's birth",
        y = "Share of director's with specific education",
        caption = "Smoothing by 3 year rolling mean"
    ) +
    scale_y_continuous(labels = scales::percent_format(scale = 100)) +
    scale_colour_manual(values = c("blue", "red"), labels = c("Technical", "Business")) 
```



